Learning Network Size while Training with ShrinkNets

### 2018-02-09 paper/ICML/evaluation.tex
edits

\begin{itemize}

@@

\item Whether the ShrinkNets approach results in network architectures that are
substantially different than the best network architecutres (in terms of
relative number of neurons per layered) identified in the literature.

@@

\end{itemize}

\noindent\textbf{Implementation: }We have implemented \textbf{switch Layers} and
the associated training procedure as a library in
pytorch~\cite{paszke2017automatic}. The layer can be freely mixed with other
popular layers such as convolutional layers, batchnorm layers, fully connected
layers, and used with all the traditional optimizers. We use our implementation
to evaluate ShrinkNets throughout the evaluation section.

@@

\subsubsection{Large Network Space: \texttt{CIFAR10}}

@@

\subsubsection{Small Network Space: \texttt{COVERTYPE}}

@@

We have demonstrated that it is possible to achieve networks with good accuracy
when using ShrinkNets both when the network space cannot be explored entirely
(\texttt{CIFAR10}) and when it can, e.g., \texttt{COVERTYPE}. The most important
result is not that ShrinkNets finds networks of good accuracy, but that those
networks are much smaller than those found by a static method. The impact of the
network size on inference time is the subject of our next evaluation goal.

@@

We have demonstrated that ShrinkNets find networks of similar or better accuracy
than static networks, and much smaller. Here we explore if the reduction in size
translates into an improvement of the inference time.

@@

For some applications, the absolute best accuracy is not something desirable if
it implies having enormous models. This observation motivates our experimental
methodology: for a given desired target accuracy, we pick the smallest network
that satisfies the accuracy constraint. For both \textit{ShrinkNets} and
\textit{Static Networks} we measure multiple metrics and compute the ratio
between the two. The metrics we are interested in are: the final model size (in
number of parameters) and inference speed on CPU and GPU for small batch sizes
($1$ input) and large batch sizes (depend on the dataset and hardware). Results
are shown on \cref{figure_CIFAR10} and \cref{figure_COVER} after the
distribution of sizes and accuracy. We limit our plot to the ($80-100\%$
accuracy range because we consider that models with lower accuracy have less
practical use).

Regarding size, the key observation is that size improvements are significant
for the \texttt{CIFAR10}. In the range of accuracies we are interested in,
improvements in size go from 4x to 40x. On the \texttt{COVERTYPE} dataset even
if the compression ratio is always above 1, it hardly exceeds 3x, except for
very high accuracy where \textit{ShrinkNets} obtained some particularly good
solutions.

We observe that unlike local sparsity compression based method, our method's
improvement on size translates directly to higher throughput at inference time.
\ra{For example, when running on CPU, blah, when running on GPU blah.}

When using a batch size of $1$ on CPU for \texttt{CIFAR10} as well as when using
a batch of $1$ on GPU, we do not observe any improvement due to the experiment
being dominated by noise. However, we still expect to find a difference in
latency if we were using a framework optimized for such goal.\ra{this is a bit
vague}

@@

\begin{figure*}[t]\centering
\begin{minipage}{2.7in}
\begin{center}
\includegraphics[width=\columnwidth]{CIFAR10_VGG_summary}
\vspace*{-5mm}
\caption{\label{figure_CIFAR10} Summary of the result of random
search over the hyper-parameters the \texttt{CIFAR10} dataset}
\end{center}
\vspace*{-4mm}
\end{minipage}
\begin{minipage}{2.7in}
\begin{center}
\includegraphics[width=\columnwidth]{COVER_FC_summary}
\vspace*{-5mm}
\caption{\label{figure_COVER} Summary of the result of random
search over the hyper-parameters the \texttt{COVERTYPE} dataset
\srm{label the x axes on the bottom plots}
}
\end{center}
\vspace*{-4mm}
\end{minipage}
\end{figure*}

>>>>

\begin{compactitem}

@@

\item Whether the ShrinkNets approach results in network architectures that are
substantially different than the best network architectures (in terms of
relative number of neurons per layered) identified in the literature.

@@

\end{compactitem}

\noindent\textbf{Implementation: }We implemented \textbf{switch Layers} and
the associated training procedure as a library in
pytorch~\cite{paszke2017automatic}. The layer can be freely mixed with other
popular layers such as convolutional layers, batchnorm layers, fully connected
layers, and used with all the traditional optimizers. We use our implementation
to evaluate ShrinkNets throughout the evaluation section.

@@

\subsubsection{Large Network Setting: \texttt{CIFAR10}}

@@

\subsubsection{Small Network Setting: \texttt{COVERTYPE}}

@@

We  demonstrated that it is possible to achieve networks with good accuracy
when using ShrinkNets both when the network space cannot be explored entirely
(\texttt{CIFAR10}) and when it can, e.g., \texttt{COVERTYPE}. The most important
result is not that ShrinkNets finds networks of good accuracy, but that those
networks are much smaller than those found by a static method. The impact of the
network size on inference time is the subject of our next evaluation goal.

@@

The previous experiment showed that ShrinkNets finds networks of similar or better accuracy
than static networks that are much smaller. We now explore if the reduction in size
translates into an improvement of the inference time.

@@

As noted in the introduction, for some applications, compact models that offer fast inference times
are as important as absolute accuracy. 
%This observation motivates our 
% the experiment approach described in the previous section 
% and shown in the top of Figure~\ref{figure_CIFAR10} 
% and~\ref{figure_COVER}:
% for a  desired target accuracy, the Pareto optimal line shows the smallest
% network
% that satisfies achieves a given accuracy. 
In this section, we study the relationship between accuracy, network size and inference time.
To do this, we select the smallest model 
that achieves a given accuracy for the both ShrinkNets and Static approach.
For each model, we measure the time to run inference with the model.  
We then compute the ratio of the network size and inference time
between ShrinkNets and Static at each accuracy level, and plot them on the bottom of 
 Figure~\ref{figure_CIFAR10} 
 and~\ref{figure_COVER}.
We limit our plots to the models with $80-100\%$
accuracy range because those are the ones that we consider to be practically useful.

 The middle plot in each figure shows the ratio of model size between ShrinkNets and Static (values $>$1 mean ShrinkNets are smaller) at different accuracy levels.  These figures show that is that size improvements are are particularly significant
for  \texttt{CIFAR10}. In the range of accuracies we are interested in,
improvements in size go from 4x to 40x. On the \texttt{COVERTYPE} dataset,
 the compression ratio is always above 1 but it rarely exceeds 3x, except for
very high accuracies where \textit{ShrinkNets} finds excellent, small
solutions.  The fact that the  \texttt{COVERTYPE} networks are not dramatically 
smaller is expected:  as the distribution at the top of Figure~\ref{figure_COVER} shows,
the static method is able to explore most of the parameter search space, so finds a range
of models that perform well at different sizes.

For speedup, we experimented with both CPUs and GPUs, and with different batch sizes, where batch size indicates the number of inputs simultaneously fed to the model for inference.
For each data set/GPU/CPU combination, we show results with batch size 1, as well as with a batch size large enough to fully utilize the hardware on each dataset and hardware configuration.  For example, for {\tt CIFAR10} on CPU, a batch size of 64 fully utilizes the CPU, whereas a GPU can execute many more models in parallel, so we use a larger batch size of 1024.  For {\tt COVERTYPE}, because the model is so much smaller, larger batches are needed to fully utilize the hardware.
Note that when using a batch size of $1$ on GPU, we do not expect to (and do not) observe any improvement because inference times are very small (typically about 10 $\mu s$), such that setup time dominates overall runtime.

 The bottom four graphs in each figure 
show the results.  Again, the {\tt CIFAR10} results show the benefit of the ShrinkNets approach
most dramatically.  On CPU, speedups range up to 6x depending on the batch size, with many models exceeding 3x speedup. In general, speedups are less than compression ratios, due to overheads in problem setup, invocation, and result generation  in Python/PyTorch.
  On GPU, the speedups are less substantial because the CUDA benchmarking utility that we use for evaluation can choose better algorithms for larger matrices which masks some of our benefit, although they are still often 1.5x--2x faster for large batch sizes.
The speedup results on {\tt COVERTYPE} are similar to those for network size:  because the networks are not much smaller, they are not much faster either.

A key takeaway of these speedup results is that, unlike local sparsity compression methods, our methods' improvement on size translates directly to higher throughput at inference time. \srm{Can we cite something showing that local sparsity does not improve performance even w/ small nets?}, at least for larger batches when hardware can be fully utilized.

@@

\begin{figure*}[t]\centering
\begin{minipage}{2.7in}
\begin{center}
\includegraphics[width=\columnwidth]{CIFAR10_VGG_summary}
\vspace*{-5mm}
\caption{\label{figure_CIFAR10} Summary of the result of random
search over the hyper-parameters the \texttt{CIFAR10} dataset}
\end{center}
\vspace*{-4mm}
\end{minipage}
\begin{minipage}{2.7in}
\begin{center}
\includegraphics[width=\columnwidth]{COVER_FC_summary}
\vspace*{-5mm}
\caption{\label{figure_COVER} Summary of the result of random
search over the hyper-parameters the \texttt{COVERTYPE} dataset
\srm{label the x axes on the bottom plots; change ``Desired accuracy`` ``Accuracy''}
}
\end{center}
\vspace*{-4mm}
\end{minipage}
\end{figure*}
